Compositional equivalences based on Open pNets

### 2020-02-05 2016-Weak-FH-Bisim/WeakBisim.tex
details in 5.7

\begin{figure}[h]
   \centerline{\includegraphics[width=12cm]{XFIG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

@@

Now let us give some details about the construction of the weak automation of the implementation pNet, obtained by application of the weak rules as explained above. Here we denote as $post_n$ the effect (= susbtitution) of the Strong transition $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

@@

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (s1\_c\gets s1\_c+1)$.
\medskip

@@

$ 3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{s1\_msg}\gets m, \texttt{s1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \forall n\ge 0$

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \forall n\ge 0$

$ 3'' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True, 
    (\texttt{m2\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
  \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \forall n\ge 0$

$ 5 = \openrule
         {\{\}, True, (\texttt{m2\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \forall n\ge 0$

$ 6 = \openrule
         {\{\}, True, (\texttt{s1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 100}
         \forall n\ge 0$

$ 7 = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \forall n\ge 0$

$ 7' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 201}
         \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \forall n\ge 1$

@@

Beforedefining a weak open automaton for the semantics of open pNets, it is necessary to state under which condition a pNet is unable to observe silent actions of its holes. In the setting of pNets this can simply be expressed as a condition on the synchronisation vectors. Precisely, the set of synchronisation vectors must contain vectors that let silent actions go through the pNet, i.e. synchronisation vectors where one hole does a $\tau$ transition, and the global visible action is a $\tau$. Additionally, no other synchronisation vector must be able to react on a silent action from a hole, i.e. if a synchronisation vector observes a $\tau$ from a hole it cannot synchronise it with another action nor emit an action that is not $\tau$. This is formalised as follows

@@

\subsection{Running example}
- In section \ref{} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol.

@@

  \TODO{Eric: this is still ABP ! rewrite completely, baesd on saturation example 2 pages before.}

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
open transition from each sequence. Here is one example of a short sequence,
starting with the open transition $OT_1$ of page \pageref{OT:SimpleProt},
then transmitting the message to the receiver, and to the environment. The two
firsts open transitionTs are internal, the last one is visible, so it constitute a
$\tau;\tau;a$ weak transition.

noindent
     $  \openrule{
      \emptyset; Pred=\True; Post=\{m_1\_m:=s_1\_m, m_1\_b:=s_1\_sb, s_2\_sb:=s_1\_sb\}
                      }
    {\ostate{s_1,m_0,m_0,r_0} \OTarrow{\underline{send(s_1\_m, s_1\_b)}} \ostate{s_2,m_1,m_0,r_0}}
    $

\medskip\noindent
     $  \openrule{
      \emptyset; Pred=[r_0\_rb=?b' \land ?m'=m_1\_m \land ?b=m_1\_b \land SV3\_m=?m' \land SV3\_b=?b'];
      \\ Post=\{r_1\_rb:=r_0\_rb; r_1\_m:=?m'; r_1\_b:=?b'      \}
                      }
    {\ostate{s_2,m_1,m_0,r_0} \OTarrow{\underline{recv(SV3\_m, SV3\_b)}} \ostate{s_2,m_0,m_0,r_1}}
    $

    \ERIC{Apres simplification:}

\medskip\noindent
     $  \openrule{
      \emptyset; Pred=[r_0\_rb=m_1\_b];
      \\ Post=\{r_1\_rb:=r_0\_rb; r_1\_m:=m_1\_m; r_1\_b:=m_1\_b      \}
                      }
    {\ostate{s_2,m_1,m_0,r_0} \OTarrow{\underline{recv(m_1\_m,m_1\_b)}} \ostate{s_2,m_0,m_0,r_1}}
    $
    \\

\medskip\noindent
     $  \openrule{
  \emptyset; Pred=\True; Post=r_2\_b:=r_1\_b; s_2\_b:=s_1\_b   \}
                      }
    {\ostate{s_2,m_0,m_0,r_1} \OTarrow{out(r_1\_m)} \ostate{s_2,m_0,m_0,r_2}}
    $
    \\

    \bigskip
    ...
    \bigskip

The resulting Weak Transition has a predicate $\phi =
\phi_1\land\phi_2\psi_1\land\phi_3\psi_2\psi_1 = \phi=[x=s_1\_b \land m=s_1\_m
  \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x
  \land v=out(r_1\_m)]$
and assignment
$\psi= \{r_2\_b:=r_0\_b\}$.

We can see here that in $\phi$ the application of assignments
$\psi_1,\psi_2,\psi_3$ have already ``eliminated'' the intermediate
state variables
$m_1\_m, b=m_1\_x, r_1\_m$, and not the intermediate input variables
$m,m',x,b$. But now $\phi$ includes enough information to eliminate them, giving
a final $\phi'= [v=out(s_1\_m)\land r_0\_b=s_1\_b]$, so the resulting
WOT is:

 \noindent
 $$ \openrule{
   \emptyset; Pred=[r_0\_b=s_1\_b]; Post= \{r_2\_b:=r_0\_b  \}
   }
   {\ostate{s_1,m_0,m_0,r_0}\OTWeakarrow{out(s_1\_m)} \ostate{s_2,m_0,m_0,r_2}}$$
%% \\
%% $\phi=[x=s_1\_b \land m=s_1\_m\\
%%   \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x\\
%%   \land v=out(r_1\_m)]\\
%% and \\
%% \psi= \{r_2\_b:=r_0\_b\}$

%% Note that

>>>>

\begin{figure}[h]
   \centerline{\includegraphics[width=15cm]{XFIG/SPSpecWeakOpen}}
  \caption{Weak Open Automaton of the Specification}
   \label{SimpleProtCounter:WeakSpecOA}
\end{figure}

@@

Now let us give some details about the construction of the weak automaton of the implementation pNet, obtained by application of the weak rules as explained above. Here we denote as $post_n$ the effect (= substitution) of the strong open transition $OT_n$ from Figure \ref{SimpleProtCounter:ImplOA}:
\smallskip

@@

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_c}\gets \texttt{s1\_c}+1)$.
\medskip

@@

$ 3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{s1\_msg}\gets m, \texttt{s1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \ \forall n\ge 0$

$ 3'' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True, 
    (\texttt{m2\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
  \ \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

$ 5 = \openrule
         {\{\}, True, (\texttt{m2\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \ \forall n\ge 0$

$ 6 = \openrule
         {\{\}, True, (\texttt{s1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

$ 7 = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 1$

@@

Before defining a weak open automaton for the semantics of open pNets,
it is necessary to state under which condition a pNet is unable to
observe silent actions of its holes. In the setting of pNets this can
simply be expressed as a condition on the synchronisation
vectors. Precisely, the set of synchronisation vectors must contain
vectors that let silent actions go through the pNet,
i.e. synchronisation vectors where one hole does a $\tau$ transition,
and the global visible action is a $\tau$. Additionally, no other
synchronisation vector must be able to react on a silent action from a
hole, i.e. if a synchronisation vector observes a $\tau$ from a hole
it cannot synchronise it with another action nor emit an action that
is not $\tau$. This is formalised as follows:

@@

\subsection{Running example}
- In section \ref{def:weakOT} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol.

@@

  \begin{itemize}
    \item[] Both pNets trivialy verify the ``non-observability''
      condition: the only vectors having $\tau$ as an action of a
      sub-net are of the form ``$< -, \tau, -> -> \tau$''.
  \end{itemize}

  Now consider the following triples:
